---
title: "Data visualization / Visualisation de données"
output:
  html_document:
    code_folding: show
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", message = FALSE)
```

## Tidy Tuesdays

Below, you will find some of the graphs (and accompanying code) I made as part of  the [R4DS](https://twitter.com/R4DScommunity) weekly visualization challenges. All of the entries I made for #TidyTuesday can be found on my [Github](https://github.com/florencevdubois/MyTidyTuesdays) page.

### 25 February 2020 

Start by loading the necessary packages and the measles data.
```{r}
# packages
library(tidyverse)
library(pacman)
library(tidyverse)

# data
measles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-25/measles.csv')
```

Let's explore the data.
```{r}
summary(measles$mmr) # distribution of vaccination rates
summary(measles$overall) # distribution of vaccination rates

weird <- measles %>% 
  filter(overall == -1 | mmr == -1)# who are the "-1" ? Apparently they're missing data.
table(weird$state) # some have missing data for "mmr", some have missing data for "overall" -- we will use this later

length(unique(measles$county)) # how many unique counties?
table(measles$year, useNA = "always") # which years?
```

Then, we need to do some data clean-up.
```{r}
measles_plot_viz <- measles %>% 
  distinct(state, name, county, overall, mmr, lat, lng, enroll) %>% # remove duplicates
  mutate(drop = ifelse(mmr >= 0 & overall >= 0, "drop", "keep")) %>% 
  filter(lng < 0) %>% # remove schools outside of the continent
  gather(type, value, overall:mmr) %>% 
  filter(value >= 0) %>% 
  filter(!(drop == "drop" & type == "overall"))
```

Finally, let's draw the map! I will be using the `maps` package to do so. I had help from Kieran Healy's [book](https://socviz.co/maps.html#maps) during this step, as well as Timo Grossenbacher's [blog](https://timogrossenbacher.ch/2016/12/beautiful-thematic-maps-with-ggplot2-only/).
```{r}
# load packages
library(maps)
library(mapproj)

# 1. load the county data
us_counties <- map_data("county") %>% 
  mutate(state_county = str_c(region, subregion, sep = "_")) 
head(us_counties)

# 2. check the distribution of vacc. rate
summary(measles_plot_viz$value)

# 3. draw a plain county map to see if it's working
p <- ggplot(mapping = aes(x = long, y = lat, group = group), data = us_counties)
p1 <- p + geom_polygon(color = "white", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  guides(fill = FALSE) +
  theme_minimal()
p1
```

Now that we know the map is working at the vaccination data makes sense, we can plot vaccination rates across the U.S. But first, for the map's caption, we need to check which type of vaccination rate is plotted in which state:
```{r results = "markup"}
table(measles_plot_viz$type, measles_plot_viz$state) 
```

Now we can plot:
```{r results = "markup", fig.height = 15, fig.width = 15}
# 4. draw map with vacc. data
p <- ggplot()
p1 <- p + 
  geom_polygon(mapping = aes(x = long, y = lat, group = group), 
               data = us_counties, 
               fill = "#f5f5f2", 
               color = "black", 
               size = 0.1) + # add county lines
  geom_point(mapping = aes(x = lng, y = lat, color = value), 
             data = subset(measles_plot_viz, value > 50), 
             size = 0.6) + # add school dots
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + # make the map more "curvy"
  theme_void() + # remove grid and axes
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, 
                                  vjust = 0.5, 
                                  size = 15),
        plot.background = element_rect(fill = "#f5f5f2", 
                                       color = NA),
        plot.caption = element_text(hjust = 0.5, 
                                    size = 10),
        plot.margin = margin(t = 1, r = 1, b = 1, l = 1, 
                             unit = "cm"),
        legend.margin = margin(t = 2, r = 2, b = 5, l = 2, 
                               unit = "mm")) + # modify plot title/caption/background/legend
  scale_color_viridis_c(option = "plasma", 
                        direction = -1, 
                        name = "Vaccination rate",
                        guide = guide_legend(direction = "horizontal",
                                             title.position = "top",
                                             title.hjust = 0.5)) + # change legend title
  labs(title = "Schools' Vaccination Rate in 31 U.S. States",
       caption = "Measles, Mumps, and Rubella (MMR) vaccination\nrate shown in FL, ID, IA, MI, NJ, NC, OK, RI, TN, WI.\nElsewhere, the overall vaccination rate is shown.") # deifne title and subtitle
p1

# save plot using ggsave(filename = "name.png", plot = last_plot(), width = 15, height = 15)
```


